Stable individual differences in strategies within, but not between, visual search tasks

### 2018-11-30 Manuscript/manuscript.tex
Draft 0.1 Finished!

A similar range of strategies, from random to near-optimal, has been found by Irons \& Leber with the Adaptive Choice Visual Search (ACVS) paradigm\cite{irons-leber2016}, designed to explore how feature-based attention is used. This paradigm involves stimuli made up of small coloured boxes (red, blue, green and a fourth colour that varies from red, through purple, to blue and back again) with numerals written inside them (See Figure \ref{fig:exampleStimuli}). The target is a defined as a red or blue box with a numeral between 1 and 4, and on each trial one target of each colour is present. The participant's task is to find one of either target as quickly as possible and report the numeral. On trials in which the fourth colour is red (or close to red), participants should search through the blue boxes and report the blue target, as there will be fewer distracters. As the fourth colour changes to through purple to blue, participants should update their strategy and search for the red target. The results showed that participants varied substantially along two key dimensions: how frequently they used the more effective target color to search (varying from chance performance to near optimal), and how often they changed between colours. Further work has shown that these differences are stable over time (between one and ten days) with test-retest correlations of around $r \approx 0.8$\cite{irons-leber2018}.

\begin{figure*}
\centering
\subfigure[][]{\includegraphics[height=3cm]{figures/split-half.png}}
\subfigure[][]{\includegraphics[height=3cm]{figures/adaptive.png}}
\subfigure[][]{\includegraphics[height=3cm]{figures/foraging.png}}
\caption{Example stimulus from the (a) \textit{split-half}, (b) \textit{adaptive choice} and (c) \textit{foraging} paradigms.}
\label{fig:exampleStimuli}
\end{figure*}

Another example of differences in search strategy comes from the foraging literature \cite{kristjansson2014,johannesson2016}. In this context, foraging is a visual search task in which there are multiple targets on each trial. Participants were asked to search through a set of items from four categories, with two categories classed as targets. In the conjunction condition (searching for red-horizontal and green-vertical line segments among red-vertical and green-horizontal distracters), most observers searched in runs, finding all the targets of one target category, and then switching and finding the targets in the other category. This strategy has previously been observers in animal foraging \cite{dawkins1971}, and suggests holding one complex target template in mind at a time is a better strategy than switching templates. However, a sub-set of observers, termed `super-foragers', were able to change between search target categories with no cost to performance. This study has been run on touch displays\cite{kristjansson2014} (observers are asked to tap on as many targets as they can as quickly as possible) and a 3D computer game (in which participants navigate a squirrel through a virtual environment picking up targets with different feature combinations)[IN PREP], with super-foragers present in both versions. While test-retest reliability has not been measured explicitly for the foraging paradigm, the task was used as a measure to assess the effect of a six day long mindfulness retreat on cognitive performance\cite{hartkamp2017}. From a re-analysis of these data, we can estimate that the test-retest reliability for the mean run length is $r \approx 0.7$ for the feature condition and $r \approx 0.88$ for the conjunction search.

Previous research has investigated the relationship between these behaviours to \hyphenation{psychometric} measures, but to date, these differences have not shown strong correlations with other attributes. In the ACVS paradigm, Irons \& Leber found no evidence of a correlation between the proportion of optimal choices made by observers and measures of visual working memory or trait compulsivity\cite{irons-leber2016}. Similarly, the differences foraging behaviour are not accounted for by working memory or inhibitory control\cite{johannesson2017}.

A common theme emerging from these studies is the observation that individual strategies vary in their degree of effectiveness or optimality. However, ``visual search'' encompasses a wide range of specific tasks, each tapping into a different aspect of behaviour (e.g. feature-based attention, information sampling). The aim of the present study is to investigate the extent to which individual differences are stable across different visual search paradigms. Are observers who use the optimal strategy in the SHLS paradigm also more optimal in the ACVS task? Does it make sense to talk about `super-searchers' who show above average performance in a range of search tasks (analogous to the 'super-recognizers' of the face-recognition literature \cite{russell2009}? Are the super-foragers consistently better or worse than more typical searchers in the other two paradigms? As a secondary question, we will measure the test-retest reliability of the differences found in the SHLS paradigm.

@@

Sample size was determined in part by a power analysis, and in part due to constraints with counter-balancing. $n = 64$ participants means we should be able to detect correlations with $r > 0.34$ with $\alpha = 0.05$, $\beta = 0.80$ between the different visual search paradigms. The sample size is therefore of sufficient size to detect relatively small correlations.

@@

The study consists of three paradigms from the visual search literature in which large individual differences have been  found\cite{nowakowska2017, irons-leber2016, kristjansson2014}. Example stimuli can be seen in Figure \ref{fig:exampleStimuli}. A brief overview of each paradigm is given below, with full details in \textit{supplementary materials}. The three tasks will be completed over two sessions, approximately one week apart. The SHLS was run in both sessions allowing us to measure test-retest reliability. The order in which participants completed the tasks was counter-balanced. There are 16 different possible orders of tasks/conditions; four participants completed each order for a total of 64.

The display was presented on a 17-inch CRT monitor with a resolutions of $1400 \times 1050$ ($n = 40$) or $1600 \times 1200$ ($n = 24$). Stimulus generation, presentation and data collection were controlled by MATLAB and the psychophysics and eyelink toolboxes \cite{brainard1997,cornelissen2002} run on a Powermac. Participants sat $\approx 47$cm from the screen.

@@

Our results are consistent with the original SHLS study\cite{nowakowska2017}: we find a large range of individual differences in search reaction time and accuracy (see Figure \ref{fig:splithalf_summary}). Furthermore, these differences are stable from between the two testing sessions, with Pearson's $r \in [0.71, 0.89]$ ($95\%$ confidence interval) for accuracy in finding hard targets. We get similar scores for the correlation in reaction times between sessions \textit{a} and \textit{b} for hard targets, ($r \in [0.54-0.81]$), easy targets ($r \in [0.52-0.80]$) and target absent trials ($r \in [0.66-0.86]$).

\begin{figure}
\centering
\subfigure[][]{\includegraphics[width=3.25cm]{../Scripts/lineseg/scratch/acc_correlation.pdf}}
\subfigure[][]{\includegraphics[width=3.25cm]{../Scripts/lineseg/scratch/rt_correlation.pdf}}
\subfigure[][]{\includegraphics[width=3.25cm]{../Scripts/lineseg/scratch/strat_corr.pdf}}
\subfigure[][]{\includegraphics[width=3.25cm]{../Scripts/lineseg/scratch/strat_compare_meanlog_rt.pdf}}
\caption{Each point represents a participant and the error-bars indicate 95\% confidence intervals. Correlation between the two sessions of the SHLS paradigm for (a)  accuracy (TP-heterogeneous trials only); (b) reaction times and (c) search strategy (TA trials only). (d) Initial search strategy correlates with reaction times in both sessions. }
\label{fig:splithalf_summary}
\end{figure}

@@

We measured an individual's strategy as the percent of plateau trials in which the individual chose the optimal target (i.e., the target with the fewest distractors: When the variable distractor was red, the optimal choice was blue, and vice versa).
The results for the ACVS were consistent with previous findings \cite{irons-leber2016,irons-leber2018}. We can clearly see from figure \ref{fig:acvs_summary} that there are individual differences in both the proportion of optimal targets reported (range $33.62\%$ - $100.00\%$, $\bar{x} = 59.15$, $s = 16.54$) and switch rate (range = $0\% - 56.68\%$, $\bar{x} = 28.50$, $s = 14.13$).

\begin{figure}
\centering
\includegraphics[width=6.5cm]{../Scripts/adaptchoice/ac_percentOpt_hist.pdf}
\caption{Histograms showing the range of individual differences found in the ACVS task.}
\label{fig:acvs_summary}
\end{figure}

@@

The main measure of interest was average run length per trial in the conjunction condition, with a run defined as a succession of one or more of the same target type, which was followed and preceded by the other target or no target. The average run length was the mean number of target selections in a run. The multiple-target foraging results were in line with previous findings \cite{kristjansson2014,johannesson2016}, with shorter run lengths for feature foraging ($\bar{x} = 3.16$, $s = 3.14$) than conjunction foraging ($\bar{x} = 11.73$, $s = 7.09$). This suggests more frequent foraging for multiple targets concurrently when those targets were defined by features than by conjunctions. Figure \ref{fig:foraging_summary} depicts the individual differences in terms of run length: twenty ``super-searchers'' were revealed, identified by short run lengths for both feature and conjunction targets.

\begin{figure}
\centering
\includegraphics[width=6.5cm]{../Scripts/foraging/fg_conjRunLength_hist.pdf}
\caption{Histogram of mean run lengths for the conjunction foraging task.}
\label{fig:foraging_summary}
\end{figure}

@@

There are many reasons why two measurements might be uncorrelated, such as range restriction or measurement noise. The test-retest correlations on each of the individual visual search tasks rules out many of these alternatives, leaving a true absence of shared variance between these tasks as a likely explanation. One might have expected reaction time to be at least modestly correlated from one search task to the next, even in the absence of correlations in strategy, as a general factor like an individual's speed-accuracy trade-off, motivation or conscientiousness might lead to better or worse overall performance across all the tasks. However, even reaction times were not highly correlated between tasks. What we did observe were correlations between our measures of strategy and reaction time within each task, suggesting our strategy metrics are predictive of search performance, and reinforcing the interpretation that an individual who is strategic in one task is not necessarily strategic in the others.

@@

Add something about learning, experience, reinforcement - Hayhoe/Land references?

Mention \cite{vogel2008}.
From \cite{proulx2011} - \textit{Vogel and colleagues have found that an individual's ability to remember a greater number of items using working memory is related to a filtering capacity in visual search that suppresses attentional capture by distracting visual information [10,11,12]. Behavioural work in visual search has extended this research to
demonstrate that working memory correlates only with top-down
visual search performance where task relevance is crucial, but not
with  bottom-up  visual  search  tasks  where  salience  in  the
environment guides attention \cite{sobel2007}}

Mention \cite{stoet2011}.

@@

We would like to thanks Anna Nowakowska, {\'A}rn Kristj{\'a}nsson and Ian Thornton for sharing data and their helpful comments and suggestions. We would like to thank Charles Rigitano and Jacqueline Von Seth who helped with data collection. We acknowledge the funding support of the James S. McDonnell Foundation (Scholar Award to ARH).

>>>>

A similar range of strategies, from random to near-optimal, has been found by Irons \& Leber with the Adaptive Choice Visual Search (ACVS) paradigm\cite{irons-leber2016}, designed to explore how feature-based attention is used. This paradigm involves stimuli made up of small coloured boxes (red, blue, green and a fourth colour that varies from red, through purple, to blue and back again) with numerals written inside them (See Figure \ref{fig:exampleStimuli}). The target is a defined as a red or blue box containing one of four numerals (e.g., 2-5), and on each trial one target of each colour is present. The participant's task is to find one of either target as quickly as possible and report the numeral. On trials in which the fourth colour is red (or close to red), participants should search through the blue boxes and report the blue target, as there will be fewer distracters. As the fourth colour changes to through purple to blue, participants should update their strategy and search for the red target. The results showed that participants varied substantially along two key dimensions: how frequently they used the more effective target color to search (varying from chance performance to near optimal), and how often they changed between colours. Further work\cite{irons-leber2018} has shown that these differences are stable over time (between one and ten days) with test-retest correlations of around $r = 0.83$ ($95\%$ confidence interval $= [0.72, 0.90]$).

\begin{figure*}
\centering
\subfigure[][]{\includegraphics[height=3cm]{figures/split-half.png}}
\subfigure[][]{\includegraphics[height=3cm]{figures/adaptive.png}}
\subfigure[][]{\includegraphics[height=3cm]{figures/foraging.png}}
\caption{Example stimulus from the (a) \textit{split-half line segment} (SHLS), (b) \textit{adaptive choice visual search} (ACVS) and (c) \textit{mouse click foraging task} (MCFT) paradigms.}
\label{fig:exampleStimuli}
\end{figure*}

Another example of differences in search strategy comes from the foraging literature \cite{kristjansson2014,johannesson2016}. In this context, foraging is a visual search task in which there are multiple targets on each trial. Participants were asked to search through a set of items from four categories, with two categories classed as targets. In the conjunction condition (searching for red-horizontal and green-vertical line segments among red-vertical and green-horizontal distracters), most observers searched in runs, finding all the targets of one target category, and then switching and finding the targets in the other category. This strategy has previously been observed in animal foraging \cite{dawkins1971}, and suggests holding one complex target template in mind at a time is a better strategy than switching templates. However, a sub-set of observers, termed `super-foragers', were able to change between search target categories with no cost to performance. This study has been run on touch displays\cite{kristjansson2014} (observers are asked to tap on as many targets as they can as quickly as possible) and a 3D computer game (in which participants navigate a squirrel through a virtual environment picking up targets with different feature combinations)[IN PREP], with super-foragers present in both versions. While test-retest reliability has not been measured explicitly for the foraging paradigm, the task was used as a measure to assess the effect of a six day long mindfulness retreat on cognitive performance\cite{hartkamp2017}. From a re-analysis of these data, we can estimate that the test-retest reliability for the mean run length is $r \approx 0.7$ for the feature condition and $r \approx 0.88$ for the conjunction search.

Previous research has investigated the relationship between these behaviours to psychometric measures, but to date, these differences have not shown strong correlations with other attributes. In the ACVS paradigm, Irons \& Leber found no evidence of a correlation between the proportion of optimal choices made by observers and measures of visual working memory; trait impulsivity\cite{irons-leber2016}; novelty seeking; need for cognition; and intolerance of uncertainty\cite{irons-leber2018}. Similarly, the differences foraging behaviour are not accounted for by working memory or inhibitory control\cite{johannesson2017}.

A common theme emerging from these studies is the observation that individual strategies vary in their degree of effectiveness or optimality. However, ``visual search'' encompasses a wide range of specific tasks, each tapping into a different aspect of behaviour (e.g. feature-based attention, information sampling). The aim of the present study is to investigate the extent to which individual differences are stable across different visual search paradigms. Are observers who use the optimal strategy in the SHLS paradigm also more optimal in the ACVS task? Does it make sense to talk about `super-searchers' who show above average performance in a range of search tasks (analogous to the 'super-recognizers' of the face-recognition literature \cite{russell2009}? Are the super-foragers consistently better or worse than more typical searchers in the other two paradigms? As a secondary question, we will measure the test-retest reliability of the differences found in the SHLS paradigm, and compare it with existing estimates of reliability for ACVS and MCFT.

@@

Sample size was determined in part by a power analysis, and in part due to constraints with counter-balancing. $n = 64$ participants means we should be able to detect correlations with $r > 0.34$ with $\alpha = 0.05$, $\beta = 0.80$ between the different visual search paradigms. The sample is therefore of sufficient size to detect relatively small correlations.

@@

The study consists of three paradigms from the visual search literature in which large individual differences have been  found\cite{nowakowska2017, irons-leber2016, kristjansson2014}. Example stimuli can be seen in Figure \ref{fig:exampleStimuli}. A brief overview of each paradigm is given below, with full details in \textit{supplementary materials}. The three tasks were completed over two sessions, approximately one week apart. The SHLS was run in both sessions allowing us to measure test-retest reliability. The order in which participants completed the tasks was counter-balanced. There are 16 different possible orders of tasks/conditions; four participants completed each order for a total of 64.

The display was presented on a 17-inch CRT monitor with a resolution of $1400 \times 1050$ ($n = 40$) or $1600 \times 1200$ ($n = 24$). Stimulus generation, presentation and data collection were controlled by MATLAB and the psychophysics and eyelink toolboxes \cite{brainard1997,cornelissen2002} run on a Powermac. Participants sat $\approx 47$cm from the screen.

@@

Our results are consistent with the original SHLS study\cite{nowakowska2017}: we find a large range of individual differences in search reaction time and accuracy (see Figure \ref{fig:splithalf_summary}). Furthermore, these differences are stable across the two sessions, with Pearson's $r \in [0.71, 0.89]$ ($95\%$ confidence interval) for accuracy in finding hard targets. We get similar scores for the correlation in reaction times between sessions \textit{a} and \textit{b} for hard targets, ($r \in [0.54-0.81]$), easy targets ($r \in [0.52-0.80]$) and target absent trials ($r \in [0.66-0.86]$).

\begin{figure}
\centering
\subfigure[][]{\includegraphics[width=3.25cm]{../Scripts/lineseg/scratch/acc_correlation.pdf}}
\subfigure[][]{\includegraphics[width=3.25cm]{../Scripts/lineseg/scratch/rt_correlation.pdf}}
\subfigure[][]{\includegraphics[width=3.25cm]{../Scripts/lineseg/scratch/strat_corr.pdf}}
\subfigure[][]{\includegraphics[width=3.25cm]{../Scripts/lineseg/scratch/strat_compare_meanlog_rt.pdf}}
\caption{Correlation between the two sessions of the SHLS paradigm for (a)  accuracy (TP-heterogeneous trials only); (b) reaction times and (c) search strategy (TA trials only). (d) Initial search strategy correlates with reaction times in both sessions. Each point represents a participant and the error-bars indicate 95\% confidence intervals.}
\label{fig:splithalf_summary}
\end{figure}

@@

We measured an individual's strategy as the percent of plateau trials in which the individual chose the optimal target (i.e., the target with the fewest distractors: When the variable distractor was red, the optimal choice was blue, and vice versa). The results for the ACVS were consistent with previous findings \cite{irons-leber2016,irons-leber2018}. We can clearly see from figure \ref{fig:acvs_mcft_summary}(a) that there are individual differences in the proportion of optimal targets reported (range $33.62\%$ - $100.00\%$, $\bar{x} = 59.15$, $s = 16.54$) and the mean ($\log_2$) reaction times (range $1.90$ - $4.80$ seconds). As with the SHLS task, the degree to which participants follow the optimal strategy is correlated with reaction times ($r \in [-0.65, -0.25]$).

\begin{figure}
\centering
\subfigure[][]{\includegraphics[width=3.25cm]{../Scripts/scratch/acvs_opt_rt}}
\subfigure[][]{\includegraphics[width=3.25cm]{../Scripts/scratch/mcft_rl_rt}}
\caption{Correlation between strategy and reaction times for (a) ACVS and (b) MCFT. Each point represents a participant.}
\label{fig:acvs_mcft_summary}
\end{figure}

@@

The main measure of interest was average run length per trial in the conjunction condition, with a run defined as a succession of one or more of the same target type, which was followed and preceded by the other target or no target. The average run length was the mean number of target selections in a run. The multiple-target foraging results were in line with previous findings \cite{kristjansson2014,johannesson2016}, with shorter run lengths for feature foraging ($\bar{x} = 3.16$, $s = 3.14$) than conjunction foraging ($\bar{x} = 11.73$, $s = 7.09$). This suggests more frequent foraging for multiple targets concurrently when those targets were defined by features than by conjunctions. Figure \ref{fig:acvs_mcft_summary}(b) depicts the individual differences in terms of run length and the correlation with reaction time ($r \in [-0.55, -0.10]$).

@@

There are many reasons why two measurements might be uncorrelated, such as range restriction or measurement noise. The test-retest correlations on each of the individual visual search tasks rule out many of these alternatives, leaving a true absence of shared variance between these tasks as a likely explanation. One might have expected reaction time to be at least modestly correlated from one search task to the next, even in the absence of correlations in strategy, as a general factor like an individual's speed-accuracy trade-off, motivation or conscientiousness might lead to better or worse overall performance across all the tasks. However, even reaction times were not highly correlated between tasks. What we did observe were correlations between our measures of strategy and reaction time within each task, suggesting our strategy metrics are predictive of search performance, and reinforcing the interpretation that an individual who is strategic in one task is not necessarily strategic in the others.

@@

We view these findings, not as a disappointing null result, but as thought-provoking and exciting. Vogel and Awh\cite{vogel2008} argued that studying individual differences in cognitive psychology (in their case, working memory) provides valuable insight to containing potential theories of the underlying cognitive mechanisms. Our results suggest that context and structure of the task also needs to be taking into account and we hope that studying how an individual's behaviour varies across different search tasks will lead to the development of a more comprehensive theory of search.

@@

We would like to thanks Anna Nowakowska, {\'A}rn Kristj{\'a}nsson and Ian Thornton for sharing data and their helpful comments and suggestions. We would like to thank Charles Rigitano and Jacqueline Von Seth who helped with data collection. We acknowledge the funding support of the James S. McDonnell Foundation (Scholar Award to ARH) and NSF BCS-1632296 to ABL.
